---
title: "Org_RANKL_UMIcounts"
author: "yejg"
date: "2017/12/26"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE,tidy = TRUE)
```

###  Library  necessary packages
```{r,message=FALSE,message=FALSE}
library(NMF)
library(rsvd)
library(Rtsne)
library(ggplot2)
library(cowplot)
library(sva)
library(igraph)
library(cccd)
library(KernSmooth)
library(beeswarm)
library(stringr)
library(reshape2)
library(formatR)
library(destiny)
source('Fxns.R')
```

###  Load data
```{r}
RANKL_UMIs<-load_data(data_name = './Extend_data/GSE92332_Org_RANKL_UMIcounts.txt.gz')
RANKL_UMIs<-RANKL_UMIs[which(unlist(apply(RANKL_UMIs,1,sum))>0),]
RANKL_tpm = data.frame(log2(1+tpm(RANKL_UMIs)))
```

###  Select variables
```{r}
v = get.variable.genes(RANKL_UMIs, min.cv2 = 100)
var.genes = as.character(rownames(v)[v$p.adj<0.05])  # select genes
```

###  Check batch effect
```{r}
RANKL.all.cells<-colnames(RANKL_UMIs)
RANKL.all.genes<-rownames(RANKL_UMIs)

RANKL.condition<-unlist(lapply(RANKL.all.cells,function(x)return(str_split(x,'_')[[1]][3])))
RANKL.cell.groups<-unlist(lapply(RANKL.all.cells,function(x)return(str_split(x,'_')[[1]][4])))
RANKL.batches<-unlist(lapply(RANKL.all.cells,function(x)return(str_split(x,'_')[[1]][1])))

table(RANKL.batches)

batch_mean_tpm = group.means(counts = RANKL_tpm, groups = RANKL.batches)
x = batch_mean_tpm[, 1]
y = batch_mean_tpm[,2]
expr.cor = round(cor(x,y),2)
smoothScatter(x, y, nrpoints=Inf, pch=16, cex=0.25, main=sprintf("Before batch correction, correlation between \ntwo illustrative batches is %s", expr.cor), xlab="All genes Batch 2, mean log2(TPM+1)", ylab="All genes Batch  1, mean log2(TPM+1)")

```

###  Remove batch effect 
     The result show that maybe the data have batch effect
```{r}
RANKL_tpm_norm = batch.normalise.comBat(counts = as.matrix(RANKL_tpm), batch.groups = RANKL.batches)
batch_mean_tpm_norm = group.means(counts =RANKL_tpm_norm, groups = RANKL.batches)
x = batch_mean_tpm_norm[, 1]
y = batch_mean_tpm_norm[,2]
expr.cor = round(cor(x,y),2)
smoothScatter(x, y, nrpoints=Inf, pch=16, cex=0.25, main=sprintf("After batch correction, correlation between \ntwo illustrative batches is %s", expr.cor), xlab="All genes Batch 2, mean log2(TPM+1)", ylab="All genes Batch  1, mean log2(TPM+1)")

```

### TSNE,PCA
```{r}
###  TSNE
RANKL.tsne.rot<-PCA_TSNE.scores(data.tpm = RANKL_tpm_norm,data.umis = RANKL_UMIs,var_genes = var.genes,
                               data_name = './Extend_data/Org_RANK',sig.pcs = TRUE,is.var.genes = TRUE)


RANKL.tsne.rot<-data.frame(RANKL.tsne.rot)
colnames(RANKL.tsne.rot)<-c('tSNE_1','tSNE_2')
pca<-read.table('./Extend_data/Org_RANK_pca_scores.txt')
```

###  Figure a
```{r}
ggplot(data = RANKL.tsne.rot[RANKL.condition=='Control',],aes(x=tSNE_1,y=tSNE_2))+geom_point()+
  scale_color_manual(values=brewer16)+scale_fill_discrete()+theme(legend.title=element_blank())+ggtitle('Org RANKL Control')


ggplot(data = RANKL.tsne.rot[RANKL.condition=="RANKL.Day3",],aes(x=tSNE_1,y=tSNE_2))+geom_point()+
  scale_color_manual(values=brewer16)+scale_fill_discrete()+theme(legend.title=element_blank())+ggtitle('Org RANKL Day3')

ggplot(data = RANKL.tsne.rot[RANKL.condition=="RANKL.Day6",],aes(x=tSNE_1,y=tSNE_2))+geom_point()+
  scale_color_manual(values=brewer16)+scale_fill_discrete()+theme(legend.title=element_blank())+ggtitle('Org RANKL Day6')

```

### Figure b

```{r}
Tnfaip2<-Genes_mean_tpm(genes = 'Tnfaip2',tpm_data = RANKL_tpm_norm,tsne_data = RANKL.tsne.rot,title = 'M-sec')

```

###   Figure c
```{r}
Gp2<-Genes_mean_tpm(genes = 'Gp2',tpm_data = RANKL_tpm_norm,tsne_data = RANKL.tsne.rot,title =NULL)

```

###   Figure d
    Expression of microfold cell marker genes in each of the organoid cell clusters. Violin plots show the distribution of expression levels (log2(TPM + 1)) for each of ten previously reported microfold cell marker genes (columns), in the cells (points) in each of 13 clusters,
including mature microfold cells (red), identified by k-nearest-neighbour clustering of the 5,434 scRNA-seq profiles from organoids

```{r,fig.height=15,fig.width=12}
library(vioplot)
RANKL.marker.genes<-c('Anxa5','Ccl20','Ccl6','Ccl9','Gp2','Marcksl1','Pglyrp1','Rac2',
                      'Spib','Tnfaip2')

Groups.color<-ifelse(RANKL.cell.groups=='M.cell','blue','red')
Violin_Genes<-function(gene,tpm.data,cell.groups,marker.genes,groups.color){
     x<-data.frame(t(tpm.data[marker.genes,]))
     dat<-data.frame(x[,gene])
     dat$Groups<-cell.groups
     colnames(dat)<-c(gene,'Groups')
     
     print(ggplot(data = dat,aes(x=dat[,2],y=dat[,1]))+geom_violin(aes(colour=Groups.color))+coord_flip()+
                     geom_jitter(shape=16, position=position_jitter(0.1),size=1)+theme(legend.title = element_blank(),legend.position = 'none',
                                                                                      axis.text.y = element_text(colour = c(rep('blue',6),'red',rep('blue',6))))+ylab(gene)+xlab(NULL))
}



RANKL.marker.tpm<-data.frame(t(RANKL_tpm_norm[RANKL.marker.genes,]))
RANKL.marker.tpm$Cells<-RANKL.cell.groups
RANKL.marker.tpm.melt<-melt(RANKL.marker.tpm)

ggplot(data=RANKL.marker.tpm.melt,aes(x=Cells,y=value))+
  geom_violin(aes(fill=value,colour=Cells),trim = FALSE)+
  geom_jitter(shape=16,size=0.2, position=position_jitter(width = 0.2))+
  facet_wrap(~ variable, nrow = 10)+
  xlab(NULL)+ylab('Expression\n(Log2(TPM+1)')+
  theme(axis.text.x = element_text(size=10,angle = 90,face = 'bold',
                                   colour = c(rep('blue',6),'red',rep('blue',6))))
```

### Figure e
####   Method 1(ggplot2)
```{r}
heatmap.cells.e<-c("Endocrine","Tuft","Enterocyte.1","Stem","Goblet","M.cell")
heatmap.genes.e<-c('Aif1','Adm','Bcl2a1b','Bcl2a1a','C1qtnf6','Fam131a','H2-M2','Msln','Padi2',
                  'Psg27','Psg25','Rassf2','Rps6kl1','Sox8','Tnfrsf11b','Tnfrsf4','Ccl9','Anxa5',
                  'Spib','Ctsh','Fabp5','Ccl20','Pglyrp1','Rac2','Gjb2','Marcksl1','Bcl2a1d','Gp2',
                  'Slc2a6','Pold1','Ubd','Mfge8','Rab32','Myadm','Frmd4b','Ncf4','Cyba','St5','Tmprss2',
                  'Epb41l2','Mmp15','Far2','Gfer','Zmat3')

heatmap.e<-Create_plot_data(genes = heatmap.genes.e,origin.data = RANKL_tpm_norm,cell_groups = RANKL.cell.groups,
                 cells = heatmap.cells.e)

p<-ggplot(heatmap.e,aes(x=Groups,y=variable))+geom_tile(aes(fill=value))+scale_fill_gradient2(mid ='grey',high='red',name='Mean\nLog2(TPM+1)')
#scale_fill_continuous(low='lightblue',high='red')+theme_bw()


p+labs(x='',y='')+scale_x_discrete(expand = c(0,0),position = 'bottom')+
  scale_y_discrete(expand = c(0,0),position = 'right')+theme(axis.text.x = element_text(face="bold", color=c("red"), size=10,hjust = 1,angle = 90),
                                                             axis.text.y = element_text(size=8),
                                                             panel.grid.major.y  = element_blank(),
                                                             panel.grid.minor.y = element_blank(),
                                                             panel.border = element_blank(),
                                                             axis.line = element_line(colour = "black"))
                                                             

```

#####  Method 2 (aheatmap)
```{r}
Heatmap.e.tpm<-Heatmap_fun(genes=heatmap.genes.e,tpm.data = RANKL_tpm_norm,condition = heatmap.cells.e,all.condition = RANKL.cell.groups)
NMF::aheatmap(Heatmap.e.tpm[[2]],Rowv = NA,Colv = NA,
              annCol = Heatmap.e.tpm[[1]],
              scale = 'none')
```


###  Figure f
####  Method 1(ggplot2)
```{r}
heatmap.cells.f<-c("Endocrine","Tuft","Enterocyte.1","Stem","Goblet","M.cell")
heatmap.genes.f<-c('Ahr','Mtf1','Relb','Mier3','Sox8','Tulp4','Onecut2','Fosl2',
                   'Irf6','Irf2','Zfp553','Nfib','Spib','Foxp4')



heatmap.f<-Create_plot_data(genes = heatmap.genes.f,origin.data = RANKL_tpm_norm,cell_groups = RANKL.cell.groups,
                            cells = heatmap.cells.f)

p<-ggplot(heatmap.f,aes(x=Groups,y=variable))+geom_tile(aes(fill=value))+scale_fill_gradient2(mid ='grey',high='red',name='Mean\nLog2(TPM+1)')
#scale_fill_continuous(low='lightblue',high='red')+theme_bw()


p+labs(x='',y='')+scale_x_discrete(expand = c(0,0),position = 'bottom')+
  scale_y_discrete(expand = c(0,0),position = 'right')+theme(axis.text.x = element_text(face="bold", color=c("red"), size=10,hjust = 1,angle = 90),
                                                             axis.text.y = element_text(size=8),
                                                             panel.grid.major.y  = element_blank(),
                                                             panel.grid.minor.y = element_blank(),
                                                             panel.border = element_blank(),
                                                             axis.line = element_line(colour = "black"))


```

####  Method 2 (aheatmap)
```{r}
Heatmap.f.tpm<-Heatmap_fun(genes=heatmap.genes.f,tpm.data = RANKL_tpm_norm,condition = heatmap.cells.f,all.condition = RANKL.cell.groups)
NMF::aheatmap(Heatmap.f.tpm[[2]],Rowv = NA,Colv = NA,
              annCol = Heatmap.f.tpm[[1]],
              scale = 'none')
```


###  Figure g
```{r}
##  Figure g
library(vioplot)
M.cells<-RANKL_tpm_norm[var.genes,RANKL.cell.groups%in%'M.cell']
All.cells<-RANKL_tpm_norm[var.genes,]

M.cells.mean<-unlist(apply(M.cells,2,mean))
All.cells.mean<-unlist(apply(All.cells,2,mean))

vioplot(All.cells.mean,M.cells.mean,names = c('All.cells','M.cells'),col = c('pink','red'))
title(ylab = 'ln vitro M cell signature\nLog2(TPM+1)')
```

