---
title: "LargeCellSort_UMIcounts"
author: "yejg"
date: "2017/12/25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE,tidy = TRUE)
```

###  Library  necessary packages
```{r,message=FALSE,message=FALSE}
library(NMF)
library(rsvd)
library(Rtsne)
library(ggplot2)
library(cowplot)
library(sva)
library(igraph)
library(cccd)
library(KernSmooth)
library(beeswarm)
library(stringr)
library(reshape2)
library(formatR)
source('Fxns.R')
```

###  Load data
```{r}
LargeCellSort_UMIs<-load_data('./Extend_data/GSE92332_LargeCellSort_UMIcounts.txt.gz') 
###  whether need this step??
gene_expr<-as.numeric(apply(LargeCellSort_UMIs,1,sum))
LargeCellSort_UMIs<-LargeCellSort_UMIs[which(gene_expr>0),]
###  

LargeCellSort_tpm<-data.frame(log2(1+tpm(LargeCellSort_UMIs)))  # transform for tpm value
```

### Select varibles
```{r}
v = get.variable.genes(LargeCellSort_UMIs, min.cv2 = 100)
var.genes = as.character(rownames(v)[v$p.adj<0.05])
```

```{r}
cell.groups<-unlist(lapply(colnames(LargeCellSort_UMIs),function(x) return(str_split(x,'_')[[1]][4])))
batches<-unlist(lapply(colnames(LargeCellSort_UMIs),function(x) return(str_split(x,'_')[[1]][3])))
mices<-unlist(lapply(colnames(LargeCellSort_UMIs),function(x) return(str_split(x,'_')[[1]][2])))

table(batches)   

```

### Check batch effect
```{r}
batch_mean_tpm = group.means(counts = LargeCellSort_tpm, groups = batches)
x = batch_mean_tpm[, 1]
y = batch_mean_tpm[,2]
expr.cor = round(cor(x,y),2)
smoothScatter(x, y, nrpoints=Inf, pch=16, cex=0.25, main=sprintf("Before batch correction, correlation between \ntwo illustrative batches is %s", expr.cor), xlab="All genes Batch 2, mean log2(TPM+1)", ylab="All genes Batch  1, mean log2(TPM+1)")

```
####  result show that there is not batch effect


###  Unsupervise cluster
#### PCA,TSNE
```{r}
tsne.rot<-PCA_TSNE.scores(data.tpm =LargeCellSort_tpm,data.umis = LargeCellSort_UMIs,var_genes = var.genes,data_name = './Extend_data/LargeCellSort')
tsne.rot<-data.frame(tsne.rot)
colnames(tsne.rot)<-c('tSNE_1','tSNE_2')
pca<-read.table('./Extend_data/LargeCellSort_pca_scores.txt')
```


###  Figure a
   Paneth cell subsets. t-SNE of 10,396 single cells
(points) was obtained using a large cell-enriched protocol (Methods),coloured by cluster annotation 
```{r}
ggplot(data = tsne.rot,aes(x=tSNE_1,y=tSNE_2,color=cell.groups))+geom_point()+
  scale_color_manual(values=brewer16)+scale_fill_discrete()+theme(legend.title=element_blank())

```


###   Figure b
   Paneth cell subset markers
```{r}
all.genes<-rownames(LargeCellSort_tpm)
cells<-c("Paneth.2","Paneth.1")
genes<-c('Defa20','Gm15308','Defa22','Defa21','Gm15315','Gm21002','Nupr1',
         'Gm15293','Pnliprp2','Itln1','Rnase1','AY761184','Gm15284','Defa23',
         'Clps','Defa17','Gm15299','Gm15292','Guca2b')

Paneth.tpm<-Heatmap_fun(genes,tpm.data = LargeCellSort_tpm,condition = cells,all.condition = cell.groups)
aheatmap(x=Paneth.tpm[[2]],Colv = NA,Rowv = NA,annCol = Paneth.tpm[[1]])
```

###  Figure c
    Two Paneth subsets reflect regional  diversity
```{r}
Regional_UMIs<-load_data("./Extend_data/GSE92332_Regional_UMIcounts.txt.gz")
Regional_tpm<-data.frame(log2(1+tpm(Regional_UMIs)))
regional.genes<-rownames(Regional_UMIs)
region.groups<-unlist(lapply(colnames(Regional_tpm),function(x)return(str_split(x,'_')[[1]][2])))

regional.cell.groups<-unlist(lapply(colnames(Regional_tpm),function(x)return(str_split(x,'_')[[1]][4])))

regional.paneth.tpm<-Heatmap_fun(genes,tpm.data = Regional_tpm[,regional.cell.groups%in%'Paneth'],condition =unique(region.groups),all.condition = region.groups[regional.cell.groups%in%'Paneth'])

aheatmap(x=regional.paneth.tpm[[2]],Colv = NA,Rowv = NA,annCol = regional.paneth.tpm[[1]])
```
   

###    Figure e
       Regional variation of intestinal stem cells.
```{r}
genes.e<-c('Eef1g','Gstm1','Rgcc','Wfdc10','Amica1','Ifitm3','Gkn3','Uba52','Acot1',
           'Nrtn','Eef1b2','Hmgcs2','Hspa8','Pdgfa','Btf3','2210407C18Rik','Clca3b','Sord',
           'Tfpi','Bex4','Bex1','Gas6','Bspry','Cyba')

regional.stem.tpm<-Heatmap_fun(genes.e,tpm.data = Regional_tpm[,regional.cell.groups%in%'Stem'],condition =unique(region.groups),all.condition = region.groups[regional.cell.groups%in%'Paneth'])
aheatmap(x=regional.stem.tpm[[2]],Colv = NA,Rowv = NA,annCol = regional.stem.tpm[[1]])
```

