---
title: "EEC_UMIcounts"
author: "yejg"
date: "2017/12/25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE,tidy = TRUE)
```

###  Library  necessary packages
```{r,message=FALSE,message=FALSE}
library(NMF)
library(rsvd)
library(Rtsne)
library(ggplot2)
library(cowplot)
library(sva)
library(igraph)
library(cccd)
library(KernSmooth)
library(beeswarm)
library(stringr)
library(reshape2)
library(formatR)
library(destiny)
source('Fxns.R')
```

###  Load data
```{r}
EEC_UMIs<-load_data('./Extend_data/GSE92332_EEC_UMIcounts.txt.gz')

###  whether need this step??
gene_expr<-as.numeric(apply(EEC_UMIs,1,sum))
EEC_UMIs<-EEC_UMIs[which(gene_expr>0),]

EEC_tpm=data.frame(log2(1+tpm(EEC_UMIs)))
```

###  Select variables
```{r}
v=get.variable.genes(EEC_UMIs,min.cv2=100)
var.genes=as.character(rownames(v)[v$p.adj<0.05])
```

### Extract messages from samples
```{r}
gene.names<-rownames(EEC_tpm)
sample.names<-colnames(EEC_tpm)


region.names<-unlist(lapply(sample.names,function(x)return(str_split(x,'_')[[1]][3])))
mices<-unlist(lapply(sample.names,function(x)return(str_split(x,'_')[[1]][2])))
cell.types<-unlist(lapply(sample.names,function(x)return(str_split(x,'_')[[1]][4])))
```

###   TSNE,PCA
```{r}
# do not known whether it has batch effect.  Ignore here.   If need to be checked,and will do it
tsne.rot<-PCA_TSNE.scores(data.tpm = EEC_tpm,data.umis=EEC_UMIs,var_genes = var.genes,data_name = './Extend_data/EEC',is.var.genes = TRUE)
tsne.rot<-as.data.frame(tsne.rot)
colnames(tsne.rot)<-c('tSNE_1','tSNE_2')
pca<-read.table('./Extend_data/EEC_pca_scores.txt')
```

###  Figure a
```{r}
ggplot(data = tsne.rot,aes(x=tSNE_1,y=tSNE_2,color=region.names))+geom_point()+
  scale_color_manual(values=brewer16)+scale_fill_discrete()+theme(legend.title=element_blank())+ggtitle('Regions')

```

### Figure b  
### do not how to plot???

###  Figure c
####  Figure c.1
```{r}
ggplot(data = tsne.rot,aes(x=tSNE_1,y=tSNE_2,color=cell.types))+geom_point()+
  scale_color_manual(values=brewer16)+scale_fill_discrete()+theme(legend.title=element_blank())+ggtitle('Cells Type')

```

```{r}
genes.1<-c('Neurog3','Sct','Tac1')
All.Facet.tsne.1<-data.frame()
for(gene in genes.1){
  All.Facet.tsne.1<-rbind(All.Facet.tsne.1,
                        Facet_wrap_fun(gene=gene,tpm.data = EEC_tpm,tsne.data = tsne.rot,condition = unique(cell.types),all.condition = cell.types))
}

ggplot(data=All.Facet.tsne.1,aes(x=tSNE_1,y=tSNE_2,colour=Gene.Mp))+geom_point()+
  facet_wrap(~ Gene, nrow = 5,ncol = 2)+
  theme(legend.title = element_text(size=10,color='blue',face='bold'),legend.position = 'right')+
  scale_color_gradient2(low='lightblue',mid='green',high='red',name='Log2\nTPM+1')

```

```{r}
genes.2<-c('Sst','Cck','Gcg','Ghrl')
All.Facet.tsne.2<-data.frame()
for(gene in genes.2){
  All.Facet.tsne.2<-rbind(All.Facet.tsne.2,
                        Facet_wrap_fun(gene=gene,tpm.data = EEC_tpm,tsne.data = tsne.rot,condition = unique(cell.types),all.condition = cell.types))
}

ggplot(data=All.Facet.tsne.2,aes(x=tSNE_1,y=tSNE_2,colour=Gene.Mp))+geom_point()+
  facet_wrap(~ Gene, nrow = 5,ncol = 2)+
  theme(legend.title = element_text(size=10,color='blue',face='bold'),legend.position = 'right')+
  scale_color_gradient2(low='lightblue',mid='green',high='red',name='Log2\nTPM+1')
```

```{r}
genes.3<-c('Glp','Nts','Reg4','Pyy')
All.Facet.tsne.3<-data.frame()
for(gene in genes.3){
  All.Facet.tsne.3<-rbind(All.Facet.tsne.3,
                        Facet_wrap_fun(gene=gene,tpm.data = EEC_tpm,tsne.data = tsne.rot,condition = unique(cell.types),all.condition = cell.types))
}

ggplot(data=All.Facet.tsne.3,aes(x=tSNE_1,y=tSNE_2,colour=Gene.Mp))+geom_point()+
  facet_wrap(~ Gene, nrow = 5,ncol = 2)+
  theme(legend.title = element_text(size=10,color='blue',face='bold'),legend.position = 'right')+
  scale_color_gradient2(low='lightblue',mid='green',high='red',name='Log2\nTPM+1')
```

###  Figure d

####  Figure d left
heatmap shows the expression of canonical gut hormone genes (rows) in each of 533 individual EEC cells (columns), coloured on the basis of their assignment to the clusters
```{r}
genes.p<-c('Neurog3','Sct','Tac1','Sst','Cck','Gcg','Ghrl','Gip','Nts','Reg4','Pyy')
#   check whether genes in sample
# for(gene in genes.p){
#   if(!gene%in%gene.names){
#     cat(sprintf('%s is not exist',gene))
#   }
# }

EEC.heatmap.d.tpm<-Heatmap_fun(genes=genes.p,tpm.data = EEC_tpm,condition = unique(cell.types),all.condition = cell.types)
aheatmap(x=EEC.heatmap.d.tpm[[2]],Colv = NA,Rowv = NA,annCol = EEC.heatmap.d.tpm[[1]])
```

####  Figure d right
heatmap shows for each cluster (columns) the percentage of cells (inset text) in which the
transcript for each hormone (rows) is detected.
```{r}
EEC_UMIs_1<-as.data.frame(t(EEC_UMIs[gene.names%in%genes.p,]))
Count<-lapply(unique(cell.types),function(x){
  x.1<-EEC_UMIs_1[cell.types%in%x,]
  x.2<-unlist(apply(x.1,2,sum))
  return(data.frame(x=round(x.2/sum(x.2),3)))
})
EEC_UMIs_prop<-as.data.frame(Count)*100
colnames(EEC_UMIs_prop)<-unique(cell.types)
#EEC_UMIs_prop<-data.frame(t(EEC_UMIs_prop)*100)


aheatmap(EEC_UMIs_prop, color = cubehelix1.16, border_color = list("cell"="white"), txt=EEC_UMIs_prop, Colv = NA, Rowv = NA )

```

